A Pilot Multicenter Cluster Randomized Trial to Compare The Effect of Trauma Life Support Training Programs on Patient and Provider Outcomes

### 2022-11-22 pilot-publication/manuscript.Rmd
Start writing discussion

Despite the widespread use of these training programs there are no
controlled trials showing that they impact patient outcomes
[@Mohammad2013; @Jayaraman2014; @Kadhum2020]. But there is level 1
evidence that these programs improve provider skills and practices
[@Ali1995; @Ali1996], and observational data suggesting that
they also improve patient outcomes [@Qin2019].

@@

# Discussion

@@

>>>>

Despite the widespread use of these training programs there are no
controlled trials showing that they impact patient outcomes
[@Mohammad2013; @Jayaraman2014; @Kadhum2020]. A systematic review
report that hospital trauma life support training is associated with a
reduction in patient mortality, based on studies that were either
observational or quasi-experimental without a control group
[@Jin2021].<!-- Note that some of the studies that Jin include state
that there was no effect of ATLS -->

@@

# Discussion
We show that conducting a cluster randomized controlled trial
comparing ATLS with PTC and standard care is feasible in this setting
provided that there are dedicated project officers to enrol
participants, collect data, and follow up participants. Missing data
rates were low for the primary outcome and many secondary outcomes, as
well as for key variables. Some variables had very high missing data
rates and may not be feasible to include in a full-scale trial.

<!-- Should we include dependency at discharge, for example using the
Modified Oxford Handicap Scale used in @Roberts2013?-->

We found that the ATLS arm had lower 30-day mortality compared to the
PTC and standard care arms.

We found that the PTC arm had lower 30-day mortality compared to the
ATLS and standard care arms.

We found that the standard care arm had lower 30-day mortality
compared to the ATLS and PTC arms.

We found that 30-day mortality was similar in the ATLS and PTC arms
but lower than the standard care arm.

We found that 30-day mortality was similar in the ATLS and standard
care arms but lower than the PTC arm.

We found that 30-day mortality was similar in the PTC and standard
care arms but lower than the ATLS arm.

Previous studies on the effect of in-hospital trauma life support
training on patient outcomes are observational or quasi-experimental
without a control group, with heterogeneous results
[@Ariyanayagam1992; @CioèPeña2016; @Petroze2014; VanHeng2008;
@Wang2010]. Studies from Trinidad and Tobago and, El Salvador, Rwanda,
and Cambodia found no significant effect on patient mortality after
implementing in-hospital trauma life support training
[@Ariyanayagam1992; @Cioèpeña2016; @Petroze2014; @Vanheng2008],
whereas one study from China on 820 patients found a significant
reduction in mortality, from 20 to 15%, after implementing ATLS
[@Wang2010].

Several randomised controlled trials show that ATLS is associated with
improved provider skills.

The primary patient participant outcome, all-cause 30-day mortality,
was missing in X% of patients. This may appear high, especially
compared to for example the CRASH-2 and REACT-2 trials, which report
missing primary outcome in less than 0.01% of patients [@Roberts2013;
@Sierink2016]. Like many other trauma trials, both CRASH-2 and REACT-2
used in-hospital mortality as their primary outcome measure, whereas
we attempted to follow up patients after discharge. The missing data
rate for in-hospital mortality was only X%, comparable to previous
trials.

Compare missing data rates.

Non-inferiority vs superiority.

@@

## Periodic suverys to residents
We did not distribute periodic surveys to the participating residents
but discussed challenges and suggestions that they had regarding the
scheduling or implementation of the training programs.
